Retrofitting Effect Handlers onto OCaml

### 2021-03-24 retro-concurrency.tex
edits

Since our primary motivation is to enable composable asynchronous I/O, let us
implement a simple lightweight thread library with support for forking new
threads and yielding control to other threads. We will then extend this library
with support for synchronously reading from channels and subsequently make it
asynchronous \textit{without changing the client code for asynchrony}. In order
to support forking and yielding threads, we declare the following effects:

@@

\begin{lstlisting}
let rec spawn f =
  match f () with
	...
  | effect (In_line ic) k -> continue k (input_line ic)
\end{lstlisting}
%
This uses OCaml standard library's |input_line| function to read a line
synchronously from the channel |ic| and resume the continuation |k| with the
resultant string. However, performing reads synchronously blocks the entire
scheduler, preventing other threads from running until the I/O is completed.

@@

Observe that this example does not resume a continuation more than once. This
also holds true for other use cases such as generators and coroutines. Hence,
our continuations are one-shot, and resuming the continuation more than once
raises an |Invalid_argument| exception. It is well-known that one-shot
continuations can be implemented efficiently~\cite{Bruggeman96}.

>>>>

Since our primary motivation is to enable composable asynchronous I/O, let us
implement a cooperative lightweight thread library with support for forking new
threads and yielding control to other threads. We will then extend this library
with support for synchronously reading from channels and subsequently make it
asynchronous \textit{without changing the client code for asynchrony}. In order
to support forking and yielding threads, we declare the following effects:

@@

\begin{lstlisting}
let rec spawn f =
  match f () with
	...
  | effect (In_line ic) k -> continue k (input_line ic)
\end{lstlisting}
%
This uses OCaml's standard |input_line| function to read a line
synchronously from the channel |ic| and resume the continuation |k| with the
resultant string. However, performing reads synchronously blocks the entire
scheduler, preventing other threads from running until the I/O is completed.

@@

This example does not resume a continuation more than once. This
also holds true for other use cases such as generators and coroutines. Hence,
our continuations are one-shot, and resuming the continuation more than once
raises an |Invalid_argument| exception. It is well-known that one-shot
continuations can be implemented efficiently~\cite{Bruggeman96}.
